---
title: "Getting Started with microscapeR"
author: "R. Eric Collins"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Getting Started with microscapeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Installation

Install microscapeR from Bioconductor:

```{r install}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("microscapeR")
```

# Overview

microscapeR provides downstream analysis tools for amplicon sequencing data
produced by DADA2. The typical workflow proceeds through these steps:

1. **Filter** the sequence table with `filterSeqtab()`
2. **Load** sample metadata with `loadMetadata()`
3. **Renormalize** by taxonomic group with `renormalize()`
4. **Build** a phylogenetic tree with `buildPhylogeny()`
5. **Ordinate** samples with `ordinateSamples()`
6. **Infer** co-occurrence networks with `sparccNetwork()`

# Quick Start

```{r quickstart}
library(microscapeR)
library(data.table)

## Assume 'seqtab' is a long-format data.table with columns:
##   sample, sequence, count

## 1. Filter
filt <- filterSeqtab(seqtab, minLength = 50, minSamples = 2,
                      minSeqs = 2, minReads = 1000)
dt <- filt$filtered

## 2. Load metadata
meta <- loadMetadata("metadata.tsv", seqtab = dt)

## 3. Renormalize by taxonomy
groups <- renormalize(dt, taxonomy_matrix)

## 4. Build phylogeny (prokaryotes only)
seqs <- unique(groups$prokaryote$sequence)
phylo <- buildPhylogeny(seqs, cpus = 4L)

## 5. Ordinate samples
ord <- ordinateSamples(groups$prokaryote, method = "tsne")

## 6. Correlation network
edges <- sparccNetwork(groups$prokaryote, minPrevalence = 0.1)
```

# Session Info

```{r sessioninfo, eval=TRUE}
sessionInfo()
```
