---
title: "Getting Started with microscapeR"
author: "R. Eric Collins"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Getting Started with microscapeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Installation

Install microscapeR from Bioconductor:

```{r install, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("microscapeR")
```

# Overview

microscapeR provides downstream analysis tools for amplicon sequencing data
produced by DADA2. The typical workflow proceeds through these steps:

1. **Filter** the sequence table with `filterSeqtab()`
2. **Load** sample metadata with `loadMetadata()`
3. **Renormalize** by taxonomic group with `renormalize()`
4. **Build** a phylogenetic tree with `buildPhylogeny()`
5. **Ordinate** samples with `ordinateSamples()`
6. **Infer** co-occurrence networks with `sparccNetwork()`

# Worked Example

This example uses small synthetic data so it runs without external files.

```{r example}
library(microscapeR)
library(data.table)

# Create a small test data.table
dt <- data.table(
    sample = rep(paste0("S", 1:6), each = 8),
    sequence = rep(paste0("SEQ", 1:8), 6),
    count = c(100, 50, 30, 10, 5, 20, 15, 25,
              80, 40, 20, 8, 3, 18, 12, 22,
              120, 60, 35, 12, 6, 24, 14, 28,
              90, 45, 25, 11, 4, 21, 13, 26,
              110, 55, 32, 9, 7, 19, 16, 23,
              95, 48, 28, 10, 5, 22, 11, 27)
)

# Filter
result <- filterSeqtab(dt, minLength = 3, minSamples = 2, minSeqs = 5,
                        minReads = 10)
result$stats

# Ordinate
coords <- ordinateSamples(result$filtered, method = "pca")
head(coords$sampleCoords)

# Network
edges <- sparccNetwork(result$filtered, minPrevalence = 0.5, minCorrelation = 0.1)
head(edges)
```

Building a phylogeny requires MAFFT to be installed on the system, so we
wrap this step in `\donttest{}`:

```{r phylogeny, eval=FALSE}
seqs <- c(
    "ATGCATGCATGCATGCATGCATGCATGCATGCATGCATGCATGCATGCATGC",
    "ATGCATGCATGCATGCATGCATGCTTGCATGCATGCATGCATGCATGCATGC",
    "ATGCATGCATGCATGCATGCATGCATGCAAGCATGCATGCATGCATGCATGC",
    "ATGCATGCATGCATGCATGCATGCATGCATGCATGCATGCTTGCATGCATGC"
)
phylo <- buildPhylogeny(seqs, cpus = 1L)
plot(phylo$tree)
```

# Quick Start (Full Workflow)

With real data, the full workflow looks like:

```{r quickstart, eval=FALSE}
library(microscapeR)
library(data.table)

## Assume 'seqtab' is a long-format data.table with columns:
##   sample, sequence, count

## 1. Filter
filt <- filterSeqtab(seqtab, minLength = 50, minSamples = 2,
                      minSeqs = 2, minReads = 1000)
dt <- filt$filtered

## 2. Load metadata
meta <- loadMetadata("metadata.tsv", seqtab = dt)

## 3. Renormalize by taxonomy
groups <- renormalize(dt, taxonomy_matrix)

## 4. Build phylogeny (prokaryotes only)
seqs <- unique(groups$prokaryote$sequence)
phylo <- buildPhylogeny(seqs, cpus = 4L)

## 5. Ordinate samples
ord <- ordinateSamples(groups$prokaryote, method = "tsne")

## 6. Correlation network
edges <- sparccNetwork(groups$prokaryote, minPrevalence = 0.1)
```

# Session Info

```{r sessioninfo, eval=TRUE}
sessionInfo()
```
